Monte Carlo Tree Search\\ Applied to Co-operative Problems \thanks{All authors are with the School of Computer Science and Electronic Engineering, University of Essex, Colchester CO4 3SQ, UK; email: {\tt \{pwillic,jwalto,dperez,sml\}@essex.ac.uk}).}

### 2015-10-14 report/response.tex

``This research problem that this paper will asses is how GGP agents perform when trying to solve a simple co-operative problem, with a focus on MCTS.''
\paragraph*{\textit{"MCTS can charge off achieving its own goals": it seems odd to describe an algorithm as "charging off".}}
Reworded.

``Unlike other domains where MCTS can solve tasks at its own pace and by itself ...''
\paragraph*{\textit{"This proves that the challenge...": no it doesn't, it indicates it.}}
Reworded to indicates.

``This indicates that the challenge of the task is high ...''
\paragraph*{\textit{"The shockingly good results...": too casual and subjective, overstating the case.}}
Reworded.

``Figure 10 shows the average score that each AI Agent achieved over all the maps. The two more powerful MCTS Agents performed the best, doing much better than the competition. Figure 11 shows a similar result, with the more powerful Agents typically completing the maps in fewer ticks. \emph{RandomController} is the only deviant here - it outperformed the GA's in score but was worse in average ticks. Comparing the AI \emph{Agents} when paired only with themselves in Figure 12 and Figure 13. The good results for \emph{RandomController} are potentially due to the fact that the 3 MCTS controllers and 2 GA algorithms all based their decisions on having Random as an accomplice.''

@@

``For the main data collection, we ran a round robin tournament of maps and Agents. Each Agent was paired with all 7 Agents (including a copy of itself) and played 47 games on each of the 7 maps.''
\paragraph*{\textit{Please cite something about (1+1)-ES}}
We have added a citation:

``The main technique used was a 1 + 1 Evolutionary Strategy \cite{t.back2000evolutionary-co}.''

\paragraph*{\textit{If all information is visible to the agents,
this problem domain seems not so interesting especially if
both agents are the same.}}
The forward model does only provide the ability to simulate an action pair(One for each agent) and get a true score based on that action pair. The forward model can be forwarded up to the tick limit or if the game is completed. There is no way to determine the presence of in-game objects other than potentially the \emph{Goal} due to its effect on score.
\paragraph*{\textit{I don't think GA is suitable for this domain.
Comparing against A* search with a simple evaluation function would be more interesting.}}
Referenced the design of the GA that shows validity in its theoretical to perform in real time general video game domains. The production of the \emph{VariGA} was to better adapt GA to the grid based world after observing the GA struggling to navigate. Whilst neither AI proved able to solve the puzzle element of the domain, this served to show differences between MCTS and the GA techniques. A* was considered to have too high a branching factor in this problem but was decided to be investigated in future work.

\paragraph*{\textit{"Doors are rendered in Black." $\rightarrow$ "Walls are rendered in Black."}}
We have corrected the typo in the definition of \emph{Walls} in the problem domain section.

### 2015-10-27 report/ceec.tex
Report

\author{Piers R. Williams, \and Joseph Walton-Rivers}

\title{Monte Carlo Tree Search Applied to Co-operative Problems}

\usepackage{graphicx}
\usepackage{glossaries}
\usepackage{verbatim}
\usepackage{array}
\usepackage{caption}
\usepackage{float}
\captionsetup{belowskip=8pt,aboveskip=4pt}
\graphicspath{{stats-processing/}{stats-processing/data/maps/}}

@@

\begin{document}
\maketitle
\begin{abstract}
This paper highlights an experiment to see how standard Monte Carlo Tree Search handles simple co-operative problems with no prior or provided knowledge. These problems are formed from a simple grid world that has a set of goals, doors and buttons as well as walls that cannot be walked through. Two agents have to each reach every goal present on the map. For a door to be open, an Agent must be present on at least one of the buttons that is linked to it.

When laid out correctly, the world requires each agent to do certain things at certain times in order to achieve the goal. With no modification to allow communication between the two Agents - Monte Carlo Tress Search performs well and performs very 'purposefully' when given enough computational time.
\end{abstract}

\section{Introduction}
This research problem assessed in this paper is how \gls{ggp} agents perform when trying to solve a simple co-operative problem without co-operative abilities, with a focus on \gls{mcts}.

\gls{ggp} is the field of writing \gls{ai} agents that can play a multitude of games without being written specifically for each one individually \cite{genesereth2005general}. \gls{ggp} in real time video games has a popular competition \cite{perez2014} run frequently.

Games that feature co-operation of some form between human players and \gls{ai} agents are commonplace. Most however feature very limited forms of co-operation that are typically scripted such as in most \gls{fps} games. Typically \gls{fps} games give the mere impression of co-operation, though any player that looks carefully at it will see the tell tale signs of scripting. Where \gls{fps} games typically excel at co-operation is in online modes that enable teams of humans to play against each other. Some games even provide squad structures and communication allowing direct command for the purpose of better co-ordination as in Battlefield 2142 \cite{battlefield2142}. \gls{rts} games also often have a small number of features designed to facilitate communication in a bid to facilitate co-operation. Two games that stand out for co-operation are Microsoft's Rise of Nations \cite{riseofnations} and Mad Doc Software's Empire Earth 2 \cite{empireEarth2}. Rise of Nations allowed a human and \gls{ai} player to operate the same set of units and buildings, though no communication was possible at all. This allowed a form of co-operation but the \gls{ai} operated to its own agenda. Empire Earth 2 allowed for humans and \gls{ai} agents to co-operate by letting plans be drawn up between them that could also be followed by both the human and \gls{ai} agent. These allowed a fairly complex set of instructions to be created, despite the simple interface.

A highly popular game that had an entire mode designed for co-operation between humans was Valve's Portal 2 \cite{portal2}. This featured human sized lab test mazes with elements that required players to work together by activating buttons, moving cubes and using intra-dimensional portals to get to the end goal. Both players were required to reach the goal in order to complete the level.

@@

\subsection{MCTS}

\gls{mcts} is a tree search algorithm originally designed by R{\'e}mi Coulom \cite{coulom2007efficient} and the major modification of \gls{uct} was defined by Kocsis and Szepesv{\'a}ri \cite{kocsis2006bandit} and improved by Kocsis et al \cite{kocsis2006improved}.

\gls{mcts} is a tree search algorithm that in games is typically applied directly to the action space. \gls{mcts} requires the ability to simulate action sequences and retrieve at least some form of reward (or lack of) from the results of these simulations. The basic steps for \gls{mcts} are:

\begin{enumerate}
\item{\emph{Selection} - Selection is the stage where the algorithm navigates the search tree, selecting optimal nodes(Based on the tree policy) until it reaches a leaf node.}
\item{\emph{Expansion} - Expansion is the stage whereby the leaf node is expanded by adding one of the remaining child states if it is not a terminal node.}
\item{\emph{Simulation} - Simulation is the stage where the algorithm forwards the model until a result is achieved. Their is usually a policy that defines how the simulation is made, with the simplest being random possible moves.}
\item{\emph{Backpropagation} - Backpropagation is the stage where the results of the simulation are propagated up the tree, so that they can influence the selection phase.}
\end{enumerate}

@@

Each \emph{Agent} is able to make a single action in each game tick. The actions are polled from each \emph{Agent} that is given the state of the game. When every \emph{Agent} has returned an action, the game is updated with these actions. The actions are implemented sequentially, if two \emph{Agents} return actions that result in occupying the same space, the \emph{Agent} with the lowest ID number will succeed while the other \emph{Agent} will fail and execute a \emph{No-Op}. The available 5 actions are:
\begin{itemize}
\item{\emph{Left} - This will move the \emph{Agent} one grid square to the left. (-1, 0)}
\item{\emph{Right} - This will move the \emph{Agent} one grid square to the right. (1, 0)}
\item{\emph{Up} - This will move the \emph{Agent} one grid square up. (0, -1)}
\item{\emph{Down} - This will move the \emph{Agent} one grid square down. (0, 1)}
\item{\emph{No-Op} - This will not move the \emph{Agent}. It will remain in the same location as the previous game state.}
\end{itemize}

@@

We feel that this creates a very interesting problem domain, due to placing each \emph{Agent} that is typically designed to solve problems by themselves in an environment where they must rely on another \emph{Agent} performing certain sequences of actions without being able to communicate that sequence or anything else.
\begin{comment}
\subsection{Initial Testing}
Initially, observations were made to see how \gls{mcts} handled the problem. At first, with a small computational budget it seemed like \gls{mcts} would perform poorly, with near random movement seeming to emerge.

This changed, when an agent (Agent 1 say) got near a \emph{Button} or \emph{Door}. When Agent 1 reached an item like this - Agent 2 would seemingly gain some purpose and typically head for the other items. When Agent 1 got on the \emph{Button}, Agent 2 would head straight through the open \emph{Door} and to the \emph{Goal}. Behaviour would then typically go back towards a more random approach, unless the Agent 1 got near the \emph{Door} - typically causing Agent 2 to head more purposefully towards the \emph{Button} that would allow Agent 1 to reach the \emph{Goal}.

With this observation made - the computational budget was increased ten-fold, and the \gls{mcts} agents performed significantly better. Typical behaviour would be near optimal with significantly less random sections of movement. This can likely be attributed to the increased rollout and \gls{uct} border depths allowing more foresight on such a small puzzle.
\end{comment}
\subsection{Tournament}
For the main data collection, we ran a round robin tournament of maps and Agents. Each Agent was paired with all 7 Agents (including a copy of itself) and played 47 games on each of the 7 maps.

\subsubsection{Maps}
The 7 maps that were included in the tournament were as follows:

\paragraph{\emph{SingleDoor}} This map is a basic map with two rooms, single door and a button either side.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{InitialState}
\caption{Single Door \protect\footnotemark}
\label{SingleDoor}
\end{figure}
\footnotetext{Displayed numbers are the ID numbers - \emph{Button} ID 0 will open \emph{Door} ID 0 but not \emph{Door} ID 1}
\paragraph{\emph{Pathfinding}} This map is the same as \emph{SingleDoor} but without the \emph{Doors} and \emph{Buttons}. This was to test the \emph{Agent's} ability to solve the simplest problem.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{level1e}
\caption{Pathfinding}
\label{Pathfinding}
\end{figure}
\paragraph{\emph{SymmetricSingleDoor}} This map was a modification of the \emph{SingleDoor} - making the map more symmetric and the action parts closer together.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{level2}
\caption{Symmetric Single Door}
\label{SymmetricSingleDoor}
\end{figure}
\paragraph{\emph{ExtendedSide}} This map was an extension of \emph{SingleDoor} with a second \emph{Goal} and some extra walls in the game. Spreading the action parts away from the door and buttons was the main purpose of this one.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{level3}
\caption{Extended Side}
\label{ExtendedSide}
\end{figure}
\paragraph{\emph{SideBySide}} This map was a modification of \emph{SingleDoor} - putting a second \emph{Goal} in and making the map partly mirrored. Each \emph{Agent} began in a separate room to each other instead of together.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{level4}
\caption{Side By Side}
\label{SideBySide}
\end{figure}
\paragraph{\emph{Airlock}} This map was a new map, designed to give wildly different roles to each AI - the top Agent would have to travel through two doors (Like an airlock) that the bottom agent had access to the linked \emph{Buttons}. Then the top \emph{Agent} could collect the reward and allow the bottom \emph{Agent} access to the \emph{Goal}.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{level5}
\caption{Airlock}
\label{Airlock}
\end{figure}
\paragraph{\emph{Butterfly}} This map was a new map - with two rooms each with a \emph{Goal} and a door. This was designed to force each \emph{Agent} to be let into the \emph{Goal} room, and out of it again.
\begin{figure}[H]
\centering
\includegraphics[scale=0.35]{level6}
\caption{Butterfly}
\label{Butterfly}
\end{figure}

\subsubsection{\gls{ai} Controllers}
A set of \gls{ai} Controllers were created in order to attempt to solve the problem domain. None of the agents have the ability to communicate with another agent and each agent has a single controller.
\paragraph{Random}
The random controller simply chooses one of the possible 5 actions. This is one of the simplest to implement and run.
\begin{comment}
\paragraph{Keyboard}
The keyboard controller takes input from the keyboard and translates that into a single Action. Once an action is polled, it is reset to the No Op action unless the keyboard is used again. 
\end{comment}
\paragraph{\gls{mcts}}
The \gls{mcts} controller is a simple implementation of \gls{uct} - with a fixed number of rollouts, \gls{uct} tree search depth limit and rollout border. The rollout border is how far in total, the forward model will be allowed to progress before the game is evaluated. No knowledge about the game is provided and the assumption is made that the other player will play randomly. The fixed rollout border makes a significant improvement in iterations per decision (from 1-3 to 500-600 in 40ms \footnote{Intel Core i5-3570, 8GB RAM, Windows 7 Enterprise 64bit}). This greatly improves the ability of \gls{mcts} to make informed decisions. The score at the end of a rollout is taken from the game state - so if \gls{mcts} does not see any player reach a goal, all branches will be equal.

For the tournament, three parameter sets were chosen for comparison and are shown in Table \ref{mctsTable}.
\begin{table}[!t]
\begin{center}
\begin{tabular}{|p{2.0cm}|p{2.0cm}|m{1.4cm}|m{1.4cm}|}
\hline
\textbf{Budget} & \textbf{Rollouts} & \textbf{UCT Search Limit} & \textbf{Rollout Border} \\
\hline
Small & 75 & 3 & 15 \\
Medium & 200 & 5 & 30 \\
High & 500 & 10 & 45 \\
\hline
\end{tabular}
\caption{Table of parameters for the 3 \gls{mcts} players}
\label{mctsTable}
\end{center}
\end{table}

\paragraph{Macro Action Genetic Algorithm}
The \gls{ga} algorithm was selected for its simplicity and had Macro Actions added in order to improve the forward search capability as well as the amount of computation time per decision available to it. The \gls{maga} used a population size of 10 and tournament selection of 3. Each candidate was a string of 15 actions - with each action performed 3 times in a row. This meant the \gls{maga} could "see" 45 ticks in the future. The design of the \gls{maga} algorithm was inspired by \textit{Perez et al} \cite{perez2013rolling}
\paragraph{Variable Macro Action Genetic Algorithm}
The Variable Macro Action \gls{ga} was created to try to solve the shortcomings of \gls{maga}. The primary premise was to allow the \gls{ga} to evolve the individual macro action lengths and the length of the overall sequence (number of macro actions). The main technique used was a 1 + 1 Evolutionary Strategy \cite{t.back2000evolutionary-co}. A 1 + 1 ES is a very simple \gls{ga} that maintains a single candidate, and in each iteration a mutation is made and if it is an improvement it is saved as the new best candidate. It is discarded if it is not an improvement. ES's are often much simpler than full \gls{ga} algorithms to create, often require much less memory and due to performing less within each iteration is able to use more of the available time budget than a full \gls{ga}. The algorithm was bounded by a number of parameters that are shown in Table \ref{vmagaTable}.
\begin{table}[!t]
\begin{center}
\begin{tabular}{| l | l | l |}
\hline
\textbf{Param} & \textbf{Controls} & \textbf{Value} \\
\hline
minNum & Number of Macro Actions & 3 \\
MaxNum & Number of Macro Actions & 10 \\
minLength & Length of Macro Actions & 1 \\
maxLength & Length of Macro Actions & 5 \\
numChance & Chance of altering Number & 0.25 \\
lengthChance & Chance of altering each length & 0.8 \\
actionChance & Chance of altering each action & 0.75 \\
\hline
\end{tabular}
\caption{Table of parameters for the Variable Macro Action Genetic Algorithm}
\label{vmagaTable}
\end{center}
\end{table}
These values need extensive tuning to realise the full potential of the algorithm and is a source of possible future work.
\section{Results}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{level1E-txt-pairs-ticks}
\caption{Average ticks taken to complete \emph{Pathfinding} when paired only with identical Agents}
\label{1epairedTicks}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{level1-txt-pairs-ticks}
\caption{Average ticks taken to complete \emph{SingleDoor} when paired only with identical Agents}
\label{1pairedTicks}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{scores-allmaps}
\caption{Average score of each AI Agent over all pairings over all the Maps}
\label{avgScoreAllMaps}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{ticks-allmaps}
\caption{Average Ticks taken for each AI Agent over all pairings over all the Maps}
\label{avgTicksAllMaps}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{scores-samepairs}
\caption{Average Scores for each AI Agent paired with itself over all the Maps}
\label{avgScorePairAllMaps}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{ticks-samepairs}
\caption{Average Ticks taken for each AI Agent paired with itself over all the Maps}
\label{avgTicksPairAllMaps}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width = \linewidth]{level5-txt-scores}
\caption{Average Score for each AI Agent paired with itself over \emph{Airlock}}
\label{avgScoreMap5}
\end{figure}

@@

\section{Discussion}
\subsection{Random Assumptions}
The \gls{mcts} and \gls{ai} algorithms are required to run simulations of the game state in order to function. Even just a single look ahead of this problem domain requires the algorithm to make a selection of what the other play would do. In all cases, the assumption of a random strategy was chosen due to the availability only of agents that could play generally. Whilst nesting these could enhance play, as explored by Tristan Cazenave \cite{cazenave2007reflexive}, it comes at a significant computational cost and must in the case of \gls{ggp} feature a very simple or Random player at the base of the stack. Random is exceedingly fast to calculate, and provides the ability for techniques like \gls{mcts} to sample how its behaviour works almost regardless of what the other agent does. This gives \gls{mcts} the ability to make moves that it believes would help a worst case scenario player - something that in this problem domain is likely to also help a more intelligent player. When \gls{mcts} is sitting on the button, an intelligent player would go through the open door and a random player would eventually go through the door. \gls{mcts} even has the option of locking the random player on the other side, forcing it to eventually complete some other task.

\subsection{MCTS}
\gls{mcts} performed very well in this problem domain.

As seen above, in Figure \ref{avgTicksPairAllMaps}, the medium and high budget implementations provided the two quickest completion times across all maps. The differences shown were significant as well as being the only two agents to complete on average in under 1000 ticks.  Figure \ref{avgScorePairAllMaps} showed that the medium and high budget implementations scored significantly better than all other agents.

>>>>

\author{Piers R. Williams,  Joseph Walton-Rivers, Diego Perez-Liebana, Simon M. Lucas}

\title{Monte Carlo Tree Search\\ Applied to Co-operative Problems \thanks{All authors are with the School of Computer Science and Electronic Engineering, University of Essex, Colchester CO4 3SQ, UK; email: {\tt \{pwillic,jwalto,dperez,sml\}@essex.ac.uk}).} }

\usepackage{graphicx}
\usepackage{glossaries}
\usepackage{verbatim}
\usepackage{array}
\usepackage{caption}
\usepackage{float}
\captionsetup{belowskip=12pt,aboveskip=4pt}
\graphicspath{{stats-processing/}{stats-processing/data/maps/}}

@@

\begin{document}
\maketitle
\begin{abstract}
This paper highlights an experiment to see how standard Monte Carlo Tree Search handles simple co-operative problems with no prior or provided knowledge. These problems are formed from a simple grid world that has a set of goals, doors and buttons as well as walls that cannot be walked through. Two agents have to reach every goal present on the map. For a door to be open, an agent must be present on at least one of the buttons that is linked to it. When laid out correctly, the world requires each agent to do certain things at certain times in order to achieve the goal. With no modification to allow communication between the two agents, Monte Carlo Tress Search performs well and very ``purposefully'' when given enough computational time.
\end{abstract}

\section{Introduction}
The research problem studied in this paper consists of how \gls{ggp} agents perform when trying to solve a simple co-operative problem without co-operative abilities, with a focus on \gls{mcts}. \gls{ggp} is the field of writing \gls{ai} agents that can play a multitude of games without being written specifically for each one individually \cite{genesereth2005general}. \gls{ggp} in real time video games has a popular competition \cite{perez2014} run frequently.

Games that feature co-operation of some form between human players and \gls{ai} agents are commonplace. Most however feature very limited forms of co-operation that are typically scripted such as in most \gls{fps} games. Typically \gls{fps} games give the mere impression of co-operation, though any player that looks carefully at it will see the tell tale signs of scripting. Where \gls{fps} games typically excel at co-operation is in online modes that enable teams of humans to play against each other. Some games even provide squad structures and communication allowing direct command for the purpose of better co-ordination as in \textit{Battlefield 2142} (EA Digital Illusions CE, 2006). \gls{rts} games also often have a small number of features designed to facilitate communication in a bid to facilitate co-operation. Two games that stand out for co-operation are \textit{Rise of Nations} (Big Huge Games, 2003) and \textit{Empire Earth II} (Mad Doc Software, 2005). Rise of Nations allowed a human and \gls{ai} player to operate the same set of units and buildings, though no communication was possible at all. This allowed a form of co-operation but the \gls{ai} operated to its own agenda. Empire Earth 2 allowed for humans and \gls{ai} agents to co-operate by letting plans be drawn up between them that could also be followed by both the human and \gls{ai} agent. These allowed a fairly complex set of instructions to be created, despite the simple interface.

A highly popular game that had an entire mode designed for co-operation between humans was \textit{Portal 2} (Valve, 2011). This featured human sized lab test mazes with elements that required players to work together by activating buttons, moving cubes and using intra-dimensional portals to get to the end goal. Both players were required to reach the goal in order to complete the level.

@@

\subsection{Monte Carlo Tree Search}

\gls{mcts} is a Tree Search algorithm originally proposed in 2006~\cite{coulom2007efficient, kocsis2006bandit, kocsis2006improved}. \gls{mcts} operates in the action space, building an asymmetric tree in memory that biases the search towards the most promising parts of the search space. \gls{mcts} estimates the theoretic value of the visited states by performing self-play from the state in the node, to the end of the game. The basic steps for \gls{mcts} are:

\begin{enumerate}
\item{\emph{Selection} - Selection is the stage where the algorithm navigates the search tree, selecting optimal nodes (based on the Tree Policy) until it reaches a leaf node.}
\item{\emph{Expansion} - Expansion is the stage whereby the leaf node is expanded by adding one of the remaining child states if it is not a terminal node (game end state).}
\item{\emph{Simulation} - Simulation is the stage where the algorithm forwards the model until a result is achieved. Their is usually a policy that defines how the simulation is made, with the simplest being random possible moves.}
\item{\emph{Backpropagation} - Backpropagation is the stage where the results of the simulation are propagated up the tree, so that they can influence the selection phase.}
\end{enumerate}

@@

Each \emph{Agent} is able to make a single action in each game tick. Each agent, after receiving the state of the game, must return a valid action. Once all \emph{Agents} have returned an action, the game is updated with these moves.

The actions are implemented sequentially: if two \emph{Agents} return actions that result in occupying the same space, the \emph{Agent} with the lowest ID number will succeed while the other \emph{Agent} will fail and execute a \emph{No-Op}. The available 5 actions are:
\begin{itemize}
\item{\emph{Left} - This will move the \emph{Agent} one grid square to the left. (-1, 0)}
\item{\emph{Right} - This will move the \emph{Agent} one grid square to the right. (1, 0)}
\item{\emph{Up} - This will move the \emph{Agent} one grid square up. (0, -1)}
\item{\emph{Down} - This will move the \emph{Agent} one grid square down. (0, 1)}
\item{\emph{No-Op} - This will not move the \emph{Agent}. It will remain in the same location as the previous game state.}
\end{itemize}

@@

We feel that this creates a very interesting problem domain, due to placing each \emph{Agent} that is typically designed to solve problems by themselves in an environment where they must rely on another \emph{Agent} performing certain sequences of actions without being able to communicate that sequence or anything else.

\begin{comment}
\subsection{Initial Testing}
Initially, observations were made to see how \gls{mcts} handled the problem. At first, with a small computational budget it seemed like \gls{mcts} would perform poorly, with near random movement seeming to emerge.

This changed, when an agent (say Agent 1 say) got near a \emph{Button} or \emph{Door}. When Agent 1 reached an item like this - Agent 2 would seemingly gain some purpose and typically head for the other items. When Agent 1 got on the \emph{Button}, Agent 2 would head straight through the open \emph{Door} and to the \emph{Goal}. Behaviour would then typically go back towards a more random approach, unless the Agent 1 got near the \emph{Door} - typically causing Agent 2 to head more purposefully towards the \emph{Button} that would allow Agent 1 to reach the \emph{Goal}.

With this observation made - the computational budget was increased ten-fold, and the \gls{mcts} agents performed significantly better. Typical behaviour would be near optimal with significantly less random sections of movement. This can likely be attributed to the increased rollout and \gls{uct} border depths allowing more foresight on such a small puzzle.
\end{comment}

\subsection{Tournament}
For the main data collection, we ran a round robin tournament of maps and Agents. Each Agent was paired with all $7$ Agents (including a copy of itself) and played $47$ games on each of the $7$ maps.

\subsubsection{Maps}
The $7$ maps that were included in the tournament were as follows:

\paragraph{\emph{SingleDoor}} This map, depicted in Figure~\ref{SingleDoor}, is a basic map with two rooms, a single door and a button on either side.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{InitialState}
\caption{Map \textit{Single Door} \protect\footnotemark}
\label{SingleDoor}
\end{figure}
\footnotetext{Displayed numbers are the ID numbers - \emph{Button} ID 0 will open \emph{Door} ID 0 but not \emph{Door} ID 1}

\paragraph{\emph{Pathfinding}} This map, depicted in Figure~\ref{Pathfinding}, is the same as \emph{SingleDoor} but without the \emph{Doors} and \emph{Buttons}. This has the aim of testing the \emph{Agent's} ability to solve the simplest problem.
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{level1e}
\caption{Map \textit{Pathfinding}}
\label{Pathfinding}
\end{figure}

\paragraph{\emph{SymmetricSingleDoor}} This map, shown in Figure~\ref{SymmetricSingleDoor}, is a modification of the \emph{SingleDoor} - making the map more symmetric and the action parts closer together.
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{level2}
\caption{Map \textit{Symmetric Single Door}}
\label{SymmetricSingleDoor}
\end{figure}

\paragraph{\emph{ExtendedSide}} This map, shown in Figure~\ref{ExtendedSide}, is an extension of \emph{SingleDoor} with a second \emph{Goal} and some extra walls in the game. Spreading the action parts away from the door and buttons is the design after this one.
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{level3}
\caption{Map \textit{Extended Side}}
\label{ExtendedSide}
\end{figure}

\paragraph{\emph{SideBySide}} This map, illustrated in Figure~\ref{SideBySide}, is a modification of \emph{SingleDoor}, including a second \emph{Goal} and making the map partly mirrored. Each \emph{Agent} begins in a separate room to each other instead of together.
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{level4}
\caption{Map \textit{Side By Side}}
\label{SideBySide}
\end{figure}

\paragraph{\emph{Airlock}} This map (see Figure~\ref{Airlock}) is designed to give wildly different roles to each AI: the top Agent would have to travel through two doors (like an airlock) that the bottom agent had access to the linked \emph{Buttons}. Then the top \emph{Agent} could collect the reward and allow the bottom \emph{Agent} access to the \emph{Goal}.
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{level5}
\caption{Map \textit{Airlock}}
\label{Airlock}
\end{figure}

\paragraph{\emph{Butterfly}} This map (see Figure~\ref{Butterfly}) is a map with two rooms each, with a \emph{Goal} and a door. This was designed to force each \emph{Agent} to be let into the \emph{Goal} room, and out of it again.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{level6}
\caption{Map \textit{Butterfly}}
\label{Butterfly}
\end{figure}

\subsubsection{\gls{ai} Controllers}
A set of \gls{ai} Controllers were created in order to attempt to solve the problem domain. None of the agents have the ability to communicate with another agent and each agent is managed by a single controller.

\paragraph{Random}
The random controller simply chooses one of the possible $5$ actions. This is one of the simplest to implement and run.

\begin{comment}
\paragraph{Keyboard}
The keyboard controller takes input from the keyboard and translates that into a single Action. Once an action is polled, it is reset to the No Op action unless the keyboard is used again. 
\end{comment}

\paragraph{\gls{mcts}}
The \gls{mcts} controller is a simple implementation of \gls{uct}, with a fixed number of rollouts, \gls{uct} tree search depth limit and rollout border. The rollout border is how far in total the forward model will be allowed to progress before the game state is evaluated. No knowledge about the game is provided and the assumption is made that the other player will play randomly. The fixed rollout border makes a significant improvement in iterations per decision (from $1$-$3$ to $500$-$600$ in $40$ms \footnote{Intel Core i5-3570, 8GB RAM, Windows 7 Enterprise 64bit}). This greatly improves the ability of \gls{mcts} to make informed decisions. The score at the end of a rollout is taken from the game state - so if \gls{mcts} does not see any player reach a goal, all branches will be equal. For the tournament, three parameter sets were chosen for comparison and are shown in Table \ref{mctsTable}.
\begin{table}[!t]
\begin{center}
\begin{tabular}{|p{2.0cm}|p{2.0cm}|m{1.4cm}|m{1.4cm}|}
\hline
\textbf{Budget} & \textbf{Rollouts} & \textbf{UCT Search Limit} & \textbf{Rollout Border} \\
\hline
Small & 75 & 3 & 15 \\
Medium & 200 & 5 & 30 \\
High & 500 & 10 & 45 \\
\hline
\end{tabular}
\caption{Table of parameters for the 3 \gls{mcts} players}
\label{mctsTable}
\end{center}
\end{table}

\paragraph{Macro Action Genetic Algorithm (GAController)}
The \gls{ga} algorithm was selected for its simplicity and had Macro Actions added in order to improve the forward search capability as well as the amount of computation time per decision available to it. The \gls{maga} used a population size of $10$ and tournament selection of $3$. Each candidate was a string of $15$ actions - with each action performed $3$ times in a row. This meant the \gls{maga} could "see" $45$ ticks in the future. The design of the \gls{maga} algorithm was inspired by \textit{Perez et al} \cite{perez2013rolling}.

\paragraph{Variable Macro Action Genetic Algorithm (VarGA)}
The Variable Macro Action \gls{ga} was created to try to solve the shortcomings of \gls{maga}. The primary premise was to allow the \gls{ga} to evolve the individual macro action lengths and the length of the overall sequence (number of macro actions). The main technique used was a $1 + 1$ Evolutionary Strategy (ES) \cite{t.back2000evolutionary-co}. A $1 + 1$ ES is a very simple \gls{ga} that maintains a single candidate. On each iteration, the mutation operator is applied to this individual, and it is saved as the new best candidate in case of an improvement in fitness (and discarded otherwise). This kind of ES is much simpler than a full \gls{ga}, requires much less memory and due to a small computational cost within each iteration, is able to use more of the available time budget.

The algorithm was bounded by a number of parameters (shown in Table \ref{vmagaTable}), and a further tuning of these is a source of possible future work in order to explore the potential of this algorithm.

\begin{table}[!t]
\begin{center}
\begin{tabular}{| l | l | l |}
\hline
\textbf{Param} & \textbf{Controls} & \textbf{Value} \\
\hline
minNum & Number of Macro Actions & 3 \\
MaxNum & Number of Macro Actions & 10 \\
minLength & Length of Macro Actions & 1 \\
maxLength & Length of Macro Actions & 5 \\
numChance & Chance of altering Number & 0.25 \\
lengthChance & Chance of altering each length & 0.8 \\
actionChance & Chance of altering each action & 0.75 \\
\hline
\end{tabular}
\caption{Table of parameters for the Variable Macro Action Genetic Algorithm}
\label{vmagaTable}
\end{center}
\end{table}

\section{Results}

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{level1E-txt-pairs-ticks}
\caption{Average ticks taken to complete \emph{Pathfinding} when paired only with identical Agents}
\label{1epairedTicks}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{level1-txt-pairs-ticks}
\caption{Average ticks taken to complete \emph{SingleDoor} when paired only with identical Agents}
\label{1pairedTicks}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{scores-allmaps}
\caption{Average score of each AI Agent over all pairings over all the Maps}
\label{avgScoreAllMaps}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{ticks-allmaps}
\caption{Average Ticks taken for each AI Agent over all pairings over all the Maps}
\label{avgTicksAllMaps}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{scores-samepairs}
\caption{Average Scores for each AI Agent paired with itself over all the Maps}
\label{avgScorePairAllMaps}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{ticks-samepairs}
\caption{Average Ticks taken for each AI Agent paired with itself over all the Maps}
\label{avgTicksPairAllMaps}
\end{figure}

@@

\begin{figure}[!t]
\centering
\includegraphics[width = \linewidth]{level5-txt-scores}
\caption{Average Score for each AI Agent paired with itself over \emph{Airlock}}
\label{avgScoreMap5}
\end{figure}

@@

\section{Discussion}
\subsection{Random Assumptions}
The \gls{mcts} and \gls{ai} algorithms are required to run simulations of the game state in order to function. Even just a single look ahead of this problem domain requires the algorithm to make a selection of what the other player would do. In all cases, the assumption of a random strategy was chosen due to the availability only of agents that could play generally. Whilst nesting these could enhance play, as explored by Tristan Cazenave \cite{cazenave2007reflexive}, it comes at a significant computational cost and must in the case of \gls{ggp} feature a very simple or Random player at the base of the stack. Random is exceedingly fast to calculate, and provides the ability for techniques like \gls{mcts} to sample how its behaviour works almost regardless of what the other agent does. This gives \gls{mcts} the ability to make moves that it believes would help a worst case scenario player - something that in this problem domain is likely to also help a more intelligent player. When \gls{mcts} is sitting on the button, an intelligent player would go through the open door and a random player would eventually go through it. \gls{mcts} even has the option of locking the random player on the other side, forcing it to eventually complete some other task.

\subsection{MCTS}
\gls{mcts} performed very well in this problem domain. As seen above, in Figure \ref{avgTicksPairAllMaps}, the medium and high budget implementations provided the two quickest completion times across all maps. The differences shown were significant as well as being the only two agents to complete on average in under $1000$ ticks.  Figure \ref{avgScorePairAllMaps} showed that the medium and high budget implementations scored significantly better than all other agents.

One possible reason for \gls{mcts} scoring so well in this problem domain is its use of statistics over hundreds and thousands of simulations to provide it with the ability to act in such a way that it handles all eventualities. Statistically,  in certain situations \gls{mcts} would only see rewards in the tree when it was situated on a \emph{Button}. This tended to cause the \gls{mcts} to travel towards \emph{Buttons}, increasing the possibility that its own simulations would cause it to be situated on the button. Eventually, most other \gls{ai} \emph{Agents} would cross through the open door.
